Automatic differentiation via off-policy particle filters

### 2023-10-10 mop/mop.tex
typo
mop-alpha

\documentclass[12p]{article}
\usepackage{fullpage}

@@

%%%%  PFILTER PSEUDOCODE
\begin{algorithm}[t!]
  \caption{\textbf{Measurement off policy (MOP) SMC}:
    \label{alg:mop-smc}
  }
  %% \KwIn{
  %%   Simulator for $f_{X_n|X_{n-1}}(x_n\given x_{n-1}\giventh\theta)$;
  %%   evaluator for $f_{Y_n|X_n}(y_n\given x_{n}\giventh\theta)$;
  %%   simulator for $f_{X_0}(x_0\giventh\theta)$;
  %%   parameter, $\theta$;
  %%   data, $y^*_{1:N}$;
  %%   number of particles, $J$.
  %% }
  %% \BlankLine
  Initialize filter particles:
  simulate ${X}_{0,j}^{F,\theta}\sim {f}_{{X}_{0}}\left(\mydot\giventh{\theta}\right)$ for $j$ in $\seq{1}{J}$\;
  Initialize relative weights: $w^{F,\theta}_{0,j}= 1$ for $j$ in $\seq{1}{J}$
  \;
  \For{$n$ in $\seq{1}{N}$}{
    Simulate for prediction:
    ${X}_{n,j}^{P,\theta}\sim {f}_{{X}_{n}|{X}_{n-1}}\big(\mydot|{X}_{n-1,j}^{F};{\theta}\big)$ for $j\ \mathrm{in}\ \seq{1}{J}$ \nllabel{alg:pfilter:step1}\;
    Evaluate measurement density:
    $g^{\theta}_{n,j}={f}_{{Y}_{n}|{X}_{n}}(y_{n}^{*}|{X}_{n,j}^{P,\theta}\giventh{\theta})$ for $j$ in $\seq{1}{J}$\;
    Update relative weights to compensate for resampling:
    $\displaystyle w^{P,\theta}_{n,j}= \frac{w^{F,\theta}_{n-1,j} \, g^{\theta}_{n,j}}{ g^{\phi}_{n,j}}$ for $j\ \mathrm{in}\ \seq{1}{J}$ \nllabel{alg:pfilter:weight:update}
    \;
    Conditional likelihood under $\phi$: 
    $L_n^{\phi} = \frac{1}{J}\sum_{m=1}^{J}g^{\phi}_{n,m}$
    \;
    Normalize weights:
    $\displaystyle \tilde{g}^{\phi}_{n,j}= \frac{g^{\phi}_{n,j}}{JL_n^{\phi}}$
    for $j\ \mathrm{in}\ \seq{1}{J}$
    \;
    Apply systematic resampling to select indices $k_{1:J}$ with $\prob\big(k_{j}=m\big) =\tilde{g}^{\phi}_{n,m}$ \nllabel{alg:pfilter:systematic}\;
    Resample:
    set ${X}_{n,j}^{F,\theta}={X}_{n,k_{j}}^{P,\theta}$ and
    $w_{n,j}^{F,\theta}= {w}_{n,k_{j}}^{P,\theta}$
    for $j$ in $\seq{1}{J}$ \nllabel{alg:pfilter:step2}
    \;
  }
  %% \KwOut{
  %%   Log likelihood estimate, $\loglikMC(\theta)=\sum_{n=1}^N\loglikMC_{n|1:n-1}$;
  %%   filter sample, $X^F_{n,1:J}$, for $n$ in $\seq{1}{N}$.
  %% }
  %% \KwCplx{$\bigO{J}$}
\end{algorithm}

@@

\item Here, the resampling rule for particle $j$ depends on $j$ (and therefore $\phi$ and $X^{P,\phi}_{n,j}$) but not on $\theta$ or  $X^{P,\theta}_{n,j}$. This does not have the standard form for a weighted particle filter, for which we can have general reweighting rules but they are usually a funciton of the location of the particle and the model parameter, $\theta$. That may be why it is harder to see that this is a properly weighted filter. Nevertheless, inspection of the algorithm shows that each particle is properly reweighted to account for its resampling probability, in step~\ref{alg:pfilter:weight:update}, so it should be properly weighted.

\item The final estimate of the likelihood is either
  \begin{equation} \label{mop:lik1}
    L(\theta) = \left(\frac{1}{J}\sum_{j=1}^J w^{F,\theta}_{N,j} \right)
    \prod_{n=1}^N L_n^\phi
  \end{equation}
  or
 \begin{equation}
   L^\prime(\theta) = \left(\frac{1}{J}\sum_{j=1}^J w^{P,\theta}_{N,j} \, g^{\theta}_{N,j}\right)
   \prod_{n=1}^{N-1} L_n^\phi
  \end{equation}
 with $L^\prime(\theta)$ presumably having slightly lower variance.

@@

  \item Taking the derivative with respect to $\theta$ at $\theta=\phi$, step~\ref{alg:pfilter:weight:update} looks very much like the stop gradient approach.

@@

\noindent {\bf Lemma 1}. {\it
Suppose that $\{(X_j,u_j),j=1,\dots,J\}$ is properly weighted for $f_X$, meaning that, if $X\sim f_X$,
\[
E[h(X)] \approx \sum_{j=1}^J \frac{w_j}{\sum_{k=1}^J u_k} h(X_j),
\]
for some appropriate formalization of $\approx$. Now, let $\{(Y_j,v_j),j=1,\dots,J\}$ be a sample drawn from $\{(X_j,u_j)\}$ where $(X_j,u_j)$ is represented, on average, $\pi_j J$ times. This could amount to binomial resampling having $J$ draws each with probability $p_j$, or systematic resampling. Suppose
\[
(Y_j,v_j) = \big(X_{a(j)},u_{a(j)}/\pi_{a(j)}\big),
\]
where $a$ is called the ancestor function. Then, $\{(Y_j,v_j),j=1,\dots,J\}$ is als properly weighted for $f_X$.
}

@@

Recursively applying Lemmas~1, 2 and~3, we obtain that 
%to step~\ref{alg:pfilter:step1}, Lemma~2 step~ {alg:pfilter:weight:update} and Lemma~3 to step~\ref{alg:pfilter:step2} we obtain that
the MOP filter is properly weighted.
Specifically, suppose inductively that $\big\{\big(X^{F,\theta}_{n-1,j},w^{F,\theta}_{n-1,j}\big)\big\}$ is properly weighted for $f_{X_{n-1}|Y_{1:n-1}}(x_{n-1}|y^*_{1:n-1};\theta)$.
Then, Lemma~2 tells us that $\big\{\big(X^{P,\theta}_{n,j},w^{P,\theta}_{n,j}\big)\big\}$ is properly weighted for $f_{X_{n}|Y_{1:n-1}}(x_{n}|y^*_{1:n-1};\theta)$.
Lemma~3 tells us that $\big\{\big(X^{P,\theta}_{n,j},w^{P,\theta}_{n,j} g^\theta_{n,j} \big)\big\}$ is therefore properly weighted for  $f_{X_{n}|Y_{1:n}}(x_{n}|y^*_{1:n};\theta)$.
Lemma~1 guarantees that the resampling rule, given by 
\[
\big(X^{F,\theta}_{n,j},w^{F,\theta}_{n,j}\big) = \big(X^{P,\theta}_{n,a(j)}, w^{P,\theta}_{n,j} g^\theta_{n,j}\big/ g^\phi_{n,j}\big),
\]
with resampling weights proportional to $g^\phi_{n,j}$, is therefore also properly weighted for $f_{X_{n}|Y_{1:n}}(x_{n}|y^*_{1:n};\theta)$.

This has addressed filtering, but not quite yet the likelihood evaluation.

A properly weighted conditional likelihood estimate from MOP (give the argument above) is
\[
f_{Y_n|Y_{1:n-1}}(y_n^*|y_{1_n-1}^*;\theta) \approx \frac{\sum_{j=1}^Jg^\theta_{n,j} w^{P,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}.
\]
We write the numerator as
\[
L^\phi_n \sum_{j=1}^J \left[ \frac{g^\theta_{n,j}}{g^\phi_{n,j}} w^{P,\theta}_{n,j}\right] \frac{g^\phi_{n,j}}{L_n^\phi}
\]
Using Lemma~1, we see this is properly estimated by
\[
L^\phi_n \sum_{j=1}^J w^{F,\theta}_{n,j}.
\]
This corresponds to
\[
L^\theta = \prod_{n=1}^N L^\phi_n \frac{\sum_{j=1}^J w^{F,\theta}_{n,j}}{\sum_{j=1}^J w^{P,\theta}_{n,j}}.
\]
Since $w^{F,\theta}_{n,j}=w^{P,\theta}_{n+1,j}$, this is a telescoping product. The remaining terms are
$\sum_{j=1}^J w^{P,\theta}_{0,j} = 1$ on the denominator and $\sum_{j=1}^J w^{F,\theta}_{N,j}$ on the numerator.
This derives the MOP estimate in (\ref{mop:lik1}).

>>>>

\documentclass[12p]{article}
\usepackage{fullpage}
\parskip 2mm

@@

\title{Automatic differentiation via off-policy particle filters}
\date{\today}
\maketitle

\begin{abstract}
Kevin has checked numerically that the measurement off policy (MOP) particle filter has a fixed-seed derivative matching Scibior \& Wood (SW). 
Here, we present the MOP algorithm in order to check that it is properly weighted.
The argument for this remains somewhat heuristic, reasoning about properly weighted samples in a similar way to Jun Liu's 2001 book.
We also introduce a MOP-$\alpha$ algorithm whose fixed-seed derivative interpolates between the basic ignore-resampling-in-derivative (IRID) PF ($\alpha=0$) and 
MOP ($\alpha=1$).
\end{abstract}

\section{The Measurement off policy (MOP) particle filter}

%%%%  MOP PFILTER PSEUDOCODE
%\begin{algorithm}[t!]
\begin{algorithm}
  \caption{\textbf{Measurement off policy (MOP) SMC}:
    \label{alg:mop-smc}
  }
  Initialize filter particles:
  simulate ${X}_{0,j}^{F,\theta}\sim {f}_{{X}_{0}}\left(\mydot\giventh{\theta}\right)$ for $j$ in $\seq{1}{J}$\;
  Initialize relative weights: $w^{F,\theta}_{0,j}= 1$ for $j$ in $\seq{1}{J}$
  \;
  \For{$n$ in $\seq{1}{N}$}{
    Simulate for prediction:
    ${X}_{n,j}^{P,\theta}\sim {f}_{{X}_{n}|{X}_{n-1}}\big(\mydot|{X}_{n-1,j}^{F};{\theta}\big)$ for $j\ \mathrm{in}\ \seq{1}{J}$ \nllabel{mop:step1}
    \;
    Prediction weights: $w_{n,j}^{P,\theta} = w_{n-1,j}^{F,\theta}$ for $j\ \mathrm{in}\ \seq{1}{J}$ 
    \;
    Evaluate measurement density:
    $g^{\theta}_{n,j}={f}_{{Y}_{n}|{X}_{n}}(y_{n}^{*}|{X}_{n,j}^{P,\theta}\giventh{\theta})$ for $j$ in $\seq{1}{J}$\;
    Before-resampling conditional likelihood: $\displaystyle L_n^{B,\theta} = \frac{\sum_{j=1}^Jg^\theta_{n,j} w^{P,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
    \;
    Conditional likelihood under $\phi$: 
    $L_n^{\phi} = \frac{1}{J}\sum_{m=1}^{J}g^{\phi}_{n,m}$
    \nllabel{mop:Lphi}
    \;
    Normalize weights:
    $\displaystyle \tilde{g}^{\phi}_{n,j}= \frac{g^{\phi}_{n,j}}{JL_n^{\phi}}$
    for $j\ \mathrm{in}\ \seq{1}{J}$
    \;
    Apply systematic resampling to select indices $k_{1:J}$ with $\prob\big(k_{j}=m\big) =\tilde{g}^{\phi}_{n,m}$ \nllabel{mop:systematic}\;
    Resample particles: ${X}_{n,j}^{F,\theta}={X}_{n,k_{j}}^{P,\theta}$
    \;
    Filter weights corrected for resampling:
    $\displaystyle w^{FC,\theta}_{n,j}= w^{P,\theta}_{n,j} \times \frac{ g^{\theta}_{n,j}}{ g^{\phi}_{n,j}}$ for $j\ \mathrm{in}\ \seq{1}{J}$ \nllabel{mop:weight:update}
    \;
    Resample filter weights:
    $w_{n,j}^{F,\theta}= {w}_{n,k_{j}}^{FC,\theta}$
    for $j$ in $\seq{1}{J}$ \nllabel{mop:step2}
    \;
    After-resampling conditional likelihood: $\displaystyle L_n^{A,\theta} = L_n^\phi \, \frac{\sum_{j=1}^J w^{F,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
    \;
  }
\end{algorithm}

@@

\item Here, the resampling rule for particle $j$ depends on $j$ (and therefore $\phi$ and $X^{P,\phi}_{n,j}$) but not on $\theta$ or  $X^{P,\theta}_{n,j}$. This does not have the standard form for a weighted particle filter, for which we can have general reweighting rules but they are usually a funciton of the location of the particle and the model parameter, $\theta$. That may be why it is harder to see that this is a properly weighted filter. Nevertheless, inspection of the algorithm shows that each particle is properly reweighted to account for its resampling probability, in step~\ref{mop:weight:update}, so it should be properly weighted.

\item The final estimate of the likelihood is either based on the after-resampling conditional likelihood estimate
  \begin{equation} \label{mop:likA}
    L^A(\theta) = \left(\frac{1}{J}\sum_{j=1}^J w^{F,\theta}_{N,j} \right)
    \prod_{n=1}^N L_n^\phi
  \end{equation}
  or the before-resampling estimate,
 \begin{equation}\label{mop:likB}
   L^B(\theta) = \prod_{n=1}^N \frac{\sum_{j=1}^J w^{P,\theta}_{N,j} \, g^{\theta}_{N,j}}{\sum_{j=1}^J w^{P,\theta}_{N,j}}.
  \end{equation}
 with $L^B(\theta)$ presumably having slightly lower variance.

@@

  \item Taking the derivative with respect to $\theta$ at $\theta=\phi$, step~\ref{mop:weight:update} looks very much like the stop gradient approach.

@@

\noindent {\bf Lemma 1}. {\it
Suppose that $\{(X_j,u_j),j=1,\dots,J\}$ is properly weighted for $f_X$, meaning that, if $X\sim f_X$,
\[
E[h(X)] \approx \sum_{j=1}^J \frac{w_j}{\sum_{k=1}^J u_k} h(X_j),
\]
for some appropriate formalization of $\approx$. Now, let $\{(Y_j,v_j),j=1,\dots,J\}$ be a sample drawn from $\{(X_j,u_j)\}$ where $(X_j,u_j)$ is represented, on average, $\pi_j J$ times. This could amount to binomial resampling having $J$ draws each with probability $p_j$, or systematic resampling. Suppose
\[
(Y_j,v_j) = \big(X_{a(j)},u_{a(j)}/\pi_{a(j)}\big),
\]
where $a$ is called the ancestor function. Then, $\{(Y_j,v_j),j=1,\dots,J\}$ is properly weighted for $f_X$.
}

@@

Recursively applying Lemmas~1, 2 and~3, we obtain that 
%to step~\ref{mop:step1}, Lemma~2 step~ {mop:weight:update} and Lemma~3 to step~\ref{mop:step2} we obtain that
the MOP filter is properly weighted.
Specifically, suppose inductively that $\big\{\big(X^{F,\theta}_{n-1,j},w^{F,\theta}_{n-1,j}\big)\big\}$ is properly weighted for $f_{X_{n-1}|Y_{1:n-1}}(x_{n-1}|y^*_{1:n-1};\theta)$.
Then, Lemma~2 tells us that $\big\{\big(X^{P,\theta}_{n,j},w^{P,\theta}_{n,j}\big)\big\}$ is properly weighted for $f_{X_{n}|Y_{1:n-1}}(x_{n}|y^*_{1:n-1};\theta)$.
Lemma~3 tells us that $\big\{\big(X^{P,\theta}_{n,j},w^{P,\theta}_{n,j} g^\theta_{n,j} \big)\big\}$ is therefore properly weighted for  $f_{X_{n}|Y_{1:n}}(x_{n}|y^*_{1:n};\theta)$.
Lemma~1 guarantees that the resampling rule, given by 
\[
\big(X^{F,\theta}_{n,j},w^{F,\theta}_{n,j}\big) = \big(X^{P,\theta}_{n,a(j)}, w^{P,\theta}_{n,j} g^\theta_{n,j}\big/ g^\phi_{n,j}\big),
\]
with resampling weights proportional to $g^\phi_{n,j}$, is therefore also properly weighted for $f_{X_{n}|Y_{1:n}}(x_{n}|y^*_{1:n};\theta)$.

This has addressed filtering, but not quite yet the likelihood evaluation. For this we use the following lemma.

\noindent {\bf Lemma 4}. {\it
  $f_{Y_n|Y_{1:n-1}}(y_n^*|y_{1_n-1}^*;\theta)$ is properly estimated by either the before-resampling estimate,
\begin{equation}\label{L1}
L_n^{B,\theta} =  \frac{\sum_{j=1}^Jg^\theta_{n,j} w^{P,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}},
\end{equation}
or by the after-resampling estimate,
\begin{equation}\label{L2}
L_n^{A,\theta} = L^\phi_n \frac{\sum_{j=1}^Jw^{F,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}.
\end{equation}
where $L^\phi_n$ comes from step~\ref{mop:Lphi}.
}

Here, (\ref{L1}) is a direct consequence of our earlier result that $\{ \big(X^{P,\theta}_{n,j},w^{P,\theta}_{n,j}\big) \}$ is properly weighted for $f_{X_{n}|Y_{1:n-1}}(x_{n}|y^*_{1:n-1};\theta)$.
To see  (\ref{L2}),
we write the numerator of (\ref{L1}) as
\[
L^\phi_n \sum_{j=1}^J \left[ \frac{g^\theta_{n,j}}{g^\phi_{n,j}} w^{P,\theta}_{n,j}\right] \frac{g^\phi_{n,j}}{L_n^\phi}
\]
Using Lemma~1, we see this is properly estimated by
\[
L^\phi_n \sum_{j=1}^J w^{F,\theta}_{n,j},
\]
from which we obtain (\ref{L2}).

Using Lemma~4, we obtain a likelihood estimate,
\[
L^{A,\theta} = \prod_{n=1}^N \left( L^\phi_n \, \frac{\sum_{j=1}^J w^{F,\theta}_{n,j}}{\sum_{j=1}^J w^{P,\theta}_{n,j}}\right).
\]
Since $w^{F,\theta}_{n,j}=w^{P,\theta}_{n+1,j}$, this is a telescoping product. The remaining terms are
$\sum_{j=1}^J w^{P,\theta}_{0,j} = J$ on the denominator and $\sum_{j=1}^J w^{F,\theta}_{N,j}$ on the numerator.
This derives the MOP estimate in (\ref{mop:likA}).

$L^{B,\theta}$ should generally be preferred, since there is no reason to include the extra variability from resampling when calculating the conditional log likelihood, but it lacks the nice telescoping product.

\section{MOP-$\alpha$}

%%%%  MOP(alpha) PFILTER PSEUDOCODE
\begin{algorithm}[t!]
  \caption{\textbf{Measurement off policy (MOP-$\alpha$) SMC}:
    \label{alg:mop-alpha}
  }
  Initialize filter particles:
  simulate ${X}_{0,j}^{F,\theta}\sim {f}_{{X}_{0}}\left(\mydot\giventh{\theta}\right)$ for $j$ in $\seq{1}{J}$\;
  Initialize relative weights: $w^{F,\theta}_{0,j}= 1$ for $j$ in $\seq{1}{J}$
  \;
  \For{$n$ in $\seq{1}{N}$}{
    Simulate for prediction:
    ${X}_{n,j}^{P,\theta}\sim {f}_{{X}_{n}|{X}_{n-1}}\big(\mydot|{X}_{n-1,j}^{F};{\theta}\big)$ for $j\ \mathrm{in}\ \seq{1}{J}$ \nllabel{mop-alpha:step1}
    \;
    Prediction weights with discounting: $w_{n,j}^{P,\theta} = \big(w_{n-1,j}^{F,\theta}\big)^\alpha$ for $j\ \mathrm{in}\ \seq{1}{J}$
    \nllabel{mop-alpha:discount}
    \;
    Evaluate measurement density:
    $g^{\theta}_{n,j}={f}_{{Y}_{n}|{X}_{n}}(y_{n}^{*}|{X}_{n,j}^{P,\theta}\giventh{\theta})$ for $j$ in $\seq{1}{J}$\;
    Before-resampling conditional likelihood: $\displaystyle L_n^{B,\theta,\alpha} = \frac{\sum_{j=1}^Jg^\theta_{n,j} w^{P,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
    \;
    Conditional likelihood under $\phi$: 
    $L_n^{\phi} = \frac{1}{J}\sum_{m=1}^{J}g^{\phi}_{n,m}$
    \nllabel{mop-alpha:Lphi}
    \;
    Normalize weights:
    $\displaystyle \tilde{g}^{\phi}_{n,j}= \frac{g^{\phi}_{n,j}}{JL_n^{\phi}}$
    for $j\ \mathrm{in}\ \seq{1}{J}$
    \;
    Apply systematic resampling to select indices $k_{1:J}$ with $\prob\big(k_{j}=m\big) =\tilde{g}^{\phi}_{n,m}$ \nllabel{mop-alpha:systematic}\;
    Resample particles: ${X}_{n,j}^{F,\theta}={X}_{n,k_{j}}^{P,\theta}$
    \;
    Filter weights corrected for resampling:
    $\displaystyle w^{FC,\theta}_{n,j}= w^{P,\theta}_{n,j} \times \frac{ g^{\theta}_{n,j}}{ g^{\phi}_{n,j}}$ for $j\ \mathrm{in}\ \seq{1}{J}$ \nllabel{mop-alpha:weight:update}
    \;
    Resample filter weights:
    $w_{n,j}^{F,\theta}= {w}_{n,k_{j}}^{FC,\theta}$
    for $j$ in $\seq{1}{J}$ \nllabel{mop-alpha:step2}
    \;
    After-resampling conditional likelihood: $\displaystyle L_n^{A,\theta,\alpha} = L_n^\phi \, \frac{\sum_{j=1}^J w^{F,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
    \;
  }
\end{algorithm}

\begin{itemize}
\item MOP has been rewritten from an earlier draft so that now the only difference between MOP and MOP-$\alpha$ is the discounting exponent $\alpha$ in step~\ref{mop-alpha:discount}.
\item When $\alpha=1$ we recover MOP.
\item When $\alpha=0$, $w^{P,\theta}_{n,j}=1$ and so $L^{B,\theta}_n$ corresponds to an equally weighted conditional likelihood, as in a regular particle filter, but with resampling carried out according to $\phi$ not $\theta$.
  When calculating a derivative with respect to $\theta$ at $\theta=\phi$, this corresponds to ignoring differentiation through the resampling, i.e., the fixed-seed derivative of the PF algorithm.
  To see this, we notice that the fixed seed derivative which ``ignores'' resampling is not really ignoring it; in an infinitesimal neighborhood of $\phi$, the resampling is constant and so the derivative is zero.
  The problem is not that the fixed-seed derivative is missing something, just that we cannot correctly move the derivative through an expectation for a discontinuous function.
\item When $\alpha < 1$, the product
\[
  L^{A,\alpha}(\theta)=\prod_{n=1}^N L^{A,\theta,\alpha}_n
\]
no longer has a telescoping cancellation. 
\end{itemize}

\section{Unresolved questions}

\begin{itemize}
\item $L^{A,\theta}_n$ and $L^{B,\theta}_n$ are asymptotically equivalent as $J$ increases, but are not identical.
  However, both are identical for a particle filter (i.e., MOP when $\theta=\phi$).
  Presumably, this could allow their fixed-seeed derivatives to be the same at $\theta=\phi$, though it is not obvious if this is true.
  In that case, we would have an alternative explanation of the telescoping identity of Scibior \& Wood.

\item It seems possible to carry out an extra numerical pass where one recalculates $L^A(\theta)$ taking advantage of the cancellation, which could make the derivative faster and/or more memory efficient, depending on what is the bottleneck. This may not be an immediate priority.

  \item The cancellation difference between  $L^{A,\theta}_n$ and $L^{B,\theta}_n$ is moot for $\alpha<1$, so there's not much reason not to use $L^{B,\theta}_n$.
